---
title: "Mapping fake IP addresses"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    favicon: img/battery-low.png
    logo: img/battery-low-small.png
---

Column {data-width=650}
-------------------------------------
    
### Interactive map
    
```{r map}
# Load packages and data --------------------------------------------------

# Install these packages once with install.packages("package_name")
# then you only need to use the library() function in future

library(readr)  # read and write files
library(dplyr)  # data manipulation and pipe operator (%>%)
library(sp)  # spatial data handling/manipulation
library(leaflet)  # leaflet mapping functions

# Read prepared data

set.seed(2412)
fake_ip_geo <- readr::read_rds("data/fake_ip_geo.RDS") %>% 
  dplyr::mutate(
    group = sample(  # randomly allocate to one of three groups
      c("Group A", "Group B", "Group C"),
      size = 500,
      replace = TRUE
    ),
    latitude = as.numeric(as.character(latitude)),  # must be numeric
    longitude = as.numeric(as.character(longitude))
  )

# Subset the data ---------------------------------------------------------

# subset by the made-up groups

group_a <- subset(fake_ip_geo, group == "Group A")
group_b <- subset(fake_ip_geo, group == "Group B")
group_c <- subset(fake_ip_geo, group == "Group C")

# Generate map ------------------------------------------------------------

# Function to build popup content based on data subsets in section above

get_popup_content <- function(group_subset) {
  paste0(
    "<b>IP: ", group_subset$ip, "</b>",
    "<br>",
    "<br>Country: ", ifelse(group_subset$country_name == "", "unknown", group_subset$country_name),
    "<br>Region: ", ifelse(group_subset$region_name == "", "unknown", group_subset$region_name),
    "<br>City: ", ifelse(group_subset$city == "", "unknown", group_subset$city),
    "<br>Time zone: ", ifelse(group_subset$time_zone == "", "unknown", group_subset$time_zone),
    "<br>",
    "<br>Arbitrary group: ", group_subset$group_name
  )
}

# Basic approach: the map is built up of a map layer and marker layers

leaflet::leaflet() %>%
  
  # add underlying map
  leaflet::addProviderTiles(providers$OpenStreetMap) %>%
  
  # group A points layer
  leaflet::addAwesomeMarkers(
    group = "Group A",  # the name of this group
    data = group_a,  # the subsetted data
    popup = ~get_popup_content(group_a),
      icon = awesomeIcons(
      library = "ion",
      icon = sample(
        c(
          "ion-battery-empty",
          "ion-battery-low",
          "ion-battery-full"
        ),
        replace = TRUE,
        nrow(group_a)
      ),
      iconColor = "white",
      markerColor = "lightred"
    )
  ) %>% 
  
  # group b points layer
  leaflet::addAwesomeMarkers(
    group = "Group B",  # the name of this group
    data = group_b,  # the subsetted data
    popup = ~get_popup_content(group_b),
            icon = awesomeIcons(
      library = "ion",
      icon = sample(
        c(
          "ion-battery-empty",
          "ion-battery-low",
          "ion-battery-full"
        ),
        replace = TRUE,
        nrow(group_b)
      ),
      iconColor = "white",
      markerColor = "red"
    )
  ) %>% 
  
  # group c points layer
  leaflet::addAwesomeMarkers(
    group = "Group C",  # the name of this group
    data = group_c,  # the subsetted data
    popup = ~get_popup_content(group_c),
    icon = awesomeIcons(
      library = "ion",
      icon = sample(
        c(
          "ion-battery-empty",
          "ion-battery-low",
          "ion-battery-full"
        ),
        replace = TRUE,
        nrow(group_c)
      ),
      iconColor = "white",
      markerColor = "darkred"
    )
  ) %>% 
  
  # add a checkbox system for turning layers on and off
  leaflet::addLayersControl(
    overlayGroups = c(
      "Group A",
      "Group B",
      "Group C"
    ),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup(c("Group B", "Group C")) %>%   # turn layers off by default
  
  # add a simple measuring tool
  leaflet::addMeasure()
```
   
Column {data-width=350 .tabset .tabset-fade}
-------------------------------------

### About

#### Purpose

To demonstrate:

* geolocation of artificially-generated [IP addresses](https://en.wikipedia.org/wiki/IP_address#Geolocation)
* plotting the coordinates of the IP addresses with arbitrarily-assigned colours and icons
* display the data in tabular form alongside the map for further interrogation
* wrapping this into a user-friendly interface

#### How to

You can:

* zoom with the + and - buttons on the map (upper left), or with your mouse wheel
* toggle marker layers with the checkboxes on the map (upper right)
* measure distances and areas with the square-rule button on the map (upper right)
* click markers to get information about that point
* search, filter and sort a table of these data in the 'Explore the data' tab

#### Key

The icons and colours are arbitrary and have no intrinsic meaning.

| Element | Meaning |
| :- | :- |
| <img src = "img/battery-full.png" style="width:50px;"> | Full battery |
| <img src = "img/battery-low.png" style="width:50px;"> | Low battery |
| <img src = "img/battery-empty.png" style="width:50px;"> | Empty battery |
| <img src = "img/group-a.png" style="width:50px;"> | Group A |
| <img src = "img/group-b.png" style="width:50px;"> | Group B |
| <img src = "img/group-c.png" style="width:50px;"> | Group C |

#### Data and tools

Coding was performed with R, using the:

* `generator` to generate fake IP addresses ([link](https://github.com/paulhendricks/generator))
* `freegeoip` to geolocate the IP addresses via API ([link](https://github.com/luiscape/freegeoip))
* `leaflet` for building the map ([link](https://rstudio.github.io/leaflet/))
* `DT` for datatables ([link](https://rstudio.github.io/DT/ package))
* `flexdashboard` for the layout ([link](https://rmarkdown.rstudio.com/flexdashboard/))
* ionicons for icons ([link](http://ionicons.com))

#### Version

Matt Dray
<br>Version 0.2
<br>Generated `r format(Sys.time(), "%d %B %Y")`

### Explore the data

```{r}
DT::datatable(
  data = fake_ip_geo[
    order(fake_ip_geo$country_name),
    -which(names(fake_ip_geo) %in% c("country_code", "region_code", "metro_code", "latitude", "longitude"))
    ],
  filter = "top",
  colnames = c("IP address", "Country name", "Region name", "City", "Zip code", "Time zone", "Group"),
  rownames = FALSE,
  options = list(pageLength = 25)
)
```